---
title: "iteration_listcols"
author: "Jasmine Zhang"
date: "2023-10-31"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(rvest)
set.seed(1)
```


### Lists
```{r}
vec_numeric = 1:4
vec_char = c("my", "name", "is", "Jasmine")
tibble(num = vec_numeric, char = vec_char)
```

Different stuff with different lengths

```{r}
l = list(
  vec_numeric = 1:5,
  vec_char = LETTERS,
  matrix = matrix(1:10, nrow = 5, ncol = 2),
  summary = summary(rnorm(100))
) #list just store anything we have

```

Accessing lists

```{r}
l$vec_char
l[[2]]
l[["summary"]]
```

### Loops

```{r}
list_norm_samples = list(
  a = rnorm(20, 1, 5),
  b = rnorm(20, 0, 7),
  c = rnorm(20, 20, 1),
  d = rnorm(20, -45, 13)
)
```

mean and sd function we wrote

```{r}
mean_and_sd = function(x){
  if(!is.numeric(x)){
    stop("Argument should be numbers")
  }else if(length(x)<2){
    stop("You need at least 2 numbers to get z scores")
  }
  mean_x = mean(x)
  sd_x = sd(x)
  
  tibble(
    mean = mean_x,
    sd = sd_x
  ) #return two values instead of the last line
}
```

```{r}
mean_and_sd(list_norm_samples$a)
mean_and_sd(list_norm_samples$b)
mean_and_sd(list_norm_samples$c)
mean_and_sd(list_norm_samples$d)
```

```{r}
output = vector("list", length = 4) #empty list to store the results we want

for (i in 1:4) {
  output[[i]] = mean_and_sd(list_norm_samples[[i]])
}

```

### Use `map`

The first argument to map is the list (or vector, or data frame) we want to iterate over, and the second argument is the function we want to apply to each element. 

```{r}
output = map(list_norm_samples, mean_and_sd) #easier to read
```

### Create df

```{r}
listcol_df = tibble(
  name = c("a", "b", "c", "d"),
  samp = list_norm_samples
)
mean_and_sd(listcol_df$samp[[1]]) #extract the first sample in the df
map(listcol_df$samp, mean_and_sd)

listcol_df |> 
  mutate(mean_sd = map(samp, mean_and_sd),
         median = map(samp, median)) |> 
  select(name, mean_sd) |> 
  unnest(mean_sd) #see exact value in tibble
```

### NSDUH
```{r}
nsduh_url = "http://samhda.s3-us-gov-west-1.amazonaws.com/s3fs-public/field-uploads/2k15StateFiles/NSDUHsaeShortTermCHG2015.htm"
nsduh_html = read_html(nsduh_url)

nsduh_table <- function(html, table_num, table_name) {
  table = 
    html |> 
    html_table() |> 
    nth(table_num) |>
    slice(-1) |> 
    select(-contains("P Value")) |>
    pivot_longer(
      -State,
      names_to = "age_year", 
      values_to = "percent") |>
    separate(age_year, into = c("age", "year"), sep = "\\(") |>
    mutate(
      year = str_replace(year, "\\)", ""),
      percent = str_replace(percent, "[a-c]$", ""),
      percent = as.numeric(percent),
      name = table_name) |>
    filter(!(State %in% c("Total U.S.", "Northeast", "Midwest", "South", "West")))
  table
}

nsduh_table(nsduh_html, 1, "marj_one_year")
```

Import data using a for loop
```{r}
table_input = list(1,4,5)
name_input = list("marj", "cocaine", "heroin")
output = vector("list", length = 3)
for (i in 1:3) {
  output[[i]] = nsduh_table(nsduh_html, table_input[[i]], name_input[[i]])
}
nsduh_df = bind_rows(output)
```

Use maps
```{r}
nsduh_import <- function(html, table_num) {
  table = 
    html |> 
    html_table() |> 
    nth(table_num) |>
    slice(-1) |> 
    select(-contains("P Value")) |>
    pivot_longer(
      -State,
      names_to = "age_year", 
      values_to = "percent") |>
    separate(age_year, into = c("age", "year"), sep = "\\(") |>
    mutate(
      year = str_replace(year, "\\)", ""),
      percent = str_replace(percent, "[a-c]$", ""),
      percent = as.numeric(percent)) |>
    filter(!(State %in% c("Total U.S.", "Northeast", "Midwest", "South", "West")))
}

nsduh_df2 = tibble(
  name = c("marj", "cocaine", "heroin"),
  number = c(1,4,5)) 
nsduh_df2 |> 
  mutate(table = map(number, nsduh_import, html = nsduh_html)) |> 
  unnest(table)

```

### revist weather dataset

```{r}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728", "USW00022534", "USS0023B17S"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2021-01-01",
    date_max = "2022-12-31") |>
  mutate(
    name = recode(
      id, 
      USW00094728 = "CentralPark_NY", 
      USW00022534 = "Molokai_HI",
      USS0023B17S = "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10) |>
  select(name, id, everything())
```

```{r}
weather_nest_df = weather_df |> 
  nest(df = date:tmin)
```
can i regress tmax on tmin for each of these?

```{r}
central_park_df = weather_nest_df |> 
  filter(name == "CentralPark_NY") |> 
  pull(df) |> 
  nth(1)

#fit a linear regression for cp
lm(tmax~tmin, central_park_df)
```

```{r}
weather_lm = function(df){
  lm(tmax~tmin, df)
}
weather_lm(central_park_df)
```
fit several df 
```{r}
#for loop
input_list = weather_nest_df |> pull(df)
output = vector("list", length = 3)
for(i in 1:3){
  output[[i]]=weather_lm(input_list[[i]])
}

weather_nest_df |> 
  mutate(models = map(df, weather_lm))
```

